---
title: 'CUNY SPS DATA607 Project2: Dataset1'
output:
  html_document:
    df_print: paged
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---
### **Name: Chinedu Onyeka, Date: October 1st, 2021**

 
#### **Objective:** To understand the trend in global child mortality from the 1980's  

#### **Data Source:** [UN Inter-agency Group for Child Mortality Estimation](https://sejdemyr.github.io/r-tutorials/basics/data/RatesDeaths_AllIndicators.xlsx)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
library(tidyverse)
```

Read the file
```{r message=FALSE}
url <- "https://raw.githubusercontent.com/chinedu2301/DATA607-Data-Acquisition-and-Management/main/dataset1_Child_Mortality.csv"
child_mort <- read_csv(url, skip = 6)
```
```{r}
head(child_mort, n = 10)
```

We don't need the ISO code and we need just the median estimate for each country
```{r}
child_mort <- child_mort %>% select(CountryName:Neonatal.Deaths.2015) %>% filter(`Uncertainty bounds*` == "Median")
head(child_mort)
```
```{r}
dim(child_mort)
```
This data-set is now 195 x 399 which is a lot of columns. More columns than rows.  

Transform from wide to long
```{r}
#Gather all Death cases into Type and Deaths.
child_mort <- child_mort %>% gather(key = "Type", value = "Deaths",
                      U5MR.1950:Neonatal.Deaths.2015, na.rm = TRUE)

head(child_mort, n = 10)
```
```{r}
#check the dimensions for this long table
dim(child_mort)
```
We now have only 5 columns and each row is for a country for each type of infant mortality and for each year.  


Separate the "Type" column:

```{r}
child_mort <- child_mort %>% separate("Type", into = c("Type", "Year")) %>%
  select(CountryName, Type:Deaths) %>% arrange(CountryName)
head(child_mort, n = 20)
```
Sub setting to data from 1990 to 2015

```{r}
child_m <- child_mort %>% filter(Year %in% 1980:2015) %>% 
  mutate(Year = as.numeric(Year))
head(child_m, n = 10)
```


```{r}
tail(child_m, n = 10)
```
Adding all categories of child mortality rates for each country,

```{r}
country_child_deaths <- child_m %>% group_by(CountryName) %>% 
  summarise(Total_Child_Deaths = sum(Deaths))
head(country_child_deaths, n = 20)
```
```{r}
country_child_deaths_year <- child_m %>% group_by(Year) %>% 
  summarise(Total_Child_Deaths = sum(Deaths))
country_child_deaths_year
```
```{r}
ggplot(data = country_child_deaths_year, aes(Year, Total_Child_Deaths)) + geom_point() +
  geom_smooth() + labs(title = "Total Child mortality vs Year") + ylab("Total Child Mortality") +
  theme_bw()
```

**Conclusion**: From the chart, we can see that the total child mortality per year around the world has significantly declined over the years since the 1980's.



