---
title: "More Practice with JSON"
author: "Coffy Andrews-Guo, Chinedu E. Onyeka"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---


Source 
[Github]()
[RPubs](https://rpubs.com/chinedu2301/823475)


## Load packages

```{r setup, message=FALSE}
library("jsonlite")
library("plyr")
library("dplyr")
library("httr")
library("purrr")
```


## Make a request to the API Server

In this assignment, The Nobel Prize, JSON data was pulled using the Developer zone API from [https://www.nobelprize.org/about/developer-zone-2/](https://www.nobelprize.org/about/developer-zone-2/). A request for data with a GET function and return a list containing all of the information returned by the API server.
```{r}
resp <- GET("https://api.nobelprize.org/2.1/laureates?_ga=2.235250305.1628049656.1634086931-1839215677.1633638535")
```

A summary provides a "Status" on the success or failure of the API request, and it comes in the form of a number. A desired number 200 corresponds to a successful request. The content type indicates what form the data takes and this response has  the data in a json format.
```{r}
resp      # summary look at the resulting response
```

Converting the raw Unicode into a character vector resembling a JSON format.
```{r results='hide'}
rawToChar(resp$content)
```

From a character vector, it is converted into list data structure using the fromJSON().The `laureates' dataset is stored in the variables consisting of three elements.
```{r}
res = fromJSON(rawToChar(resp$content))
names(res)
```

## Parse and generate data

This analysis was developed by individually extracting list elements from the 'laureates' data frame using '[ ]' and  tidyverse function 

```{r}
res1 <- as.data.frame(res$laureates$id)   # column name 'Id'
colnames(res1)[1] = "Laureates.Id"  # change column name by colname function
head(res1, 4)
```


```{r}
res2 <- as.data.frame(res$laureates$knownName)   # knownName
colnames(res2)[1] = "KnownName"  # change column name by colname function
res2 <- subset(res2, select = -c(se)) # drop columns 'se'

head(res2, 4)
```


```{r}
res3 <- as.data.frame(res$laureates$givenName)   # givenName
colnames(res3)[1] = "GivenName"  # change column name by colname function
res3 <- subset(res3, select = -c(se)) # drop columns 'se'

head(res3, 4)
```



```{r}
res4 <- as.data.frame(res$laureates$familyName)   # familyName
colnames(res4)[1] = "FamilyName"  # change column name by colname function
res4 <- subset(res4, select = -c(se)) # drop columns 'se'

head(res4, 4)
```


```{r}
res5 <- as.data.frame(res$laureates$fullName)   # fullName
colnames(res5)[1] = "FullName"  # change column name by colname function
res5 <- subset(res5, select = -c(se)) # drop columns 'se'

head(res5, 4)
```


```{r}
res6 <- as.data.frame(res$laureates$fileName)   # fileName
colnames(res6)[1] = "FileName"  # change column name by colname function

head(res6, 4)
```


```{r}
res7 <- as.data.frame(res$laureates$gender)   # gender
colnames(res7)[1] = "Gender"  # change column name by colname function

head(res7, 4)
```


```{r}
res8 <- as.data.frame(res$laureates$birth)   # birth.date
colnames(res8)[1] = "BirthDate"  # change column name by colname function
res8 <- subset(res8, select = -c(place)) # drop columns 'place'

head(res8, 4)
```


```{r}
res8.1 <- as.data.frame(res$laureates$birth$place$city)   # birthplacecity
colnames(res8.1)[1] = "Birth.PlaceCity"  # change column name by colname function
res8.1 <- subset(res8.1, select = -c(no, se)) # drop columns 'no' and 'se'

head(res8.1, 4)
```


```{r}
res8.2 <- as.data.frame(res$laureates$birth$place$country)   # birthplacecountry
colnames(res8.2)[1] = "Birth.PlaceCountry"  # change column name by colname function
res8.2 <- subset(res8.2, select = -c(no, se)) # drop columns 'no' and 'se'

head(res8.2, 4)
```


```{r}
res8.3 <- as.data.frame(res$laureates$birth$place$cityNow)   # birthplacecityNow
colnames(res8.3)[1] = "Birth.PlaceCityNow"  # change column name by colname function
res8.3 <- subset(res8.3, select = -c(no, se, sameAs)) # drop columns 'no', 'se', 'sameAs'

head(res8.3, 4)
```


```{r}
res8.4 <- as.data.frame(res$laureates$birth$place$countryNow)   # birthplacecountryNow
colnames(res8.4)[1] = "Birth.PlaceCountryNow"  # change column name by colname function
res8.4 <- subset(res8.4, select = -c(no, se, sameAs)) # drop columns 'no', 'se', and 'sameAs'

head(res8.4, 4)
```


```{r}
res8.5 <- as.data.frame(res$laureates$birth$place$continent)   # birthplacecontinent
colnames(res8.5)[1] = "Birth.PlaceContinent"  # change column name by colname function
res8.5 <- subset(res8.5, select = -c(no, se)) # drop columns 'no' and 'se'

head(res8.5, 4)
```


```{r}
res8.6 <- as.data.frame(res$laureates$birth$place$locationString)   # birthplacelocationString
colnames(res8.6)[1] = "Birth.PlaceLocationString"  # change column name by colname function
res8.6 <- subset(res8.6, select = -c(no, se)) # drop columns 'no' and 'se'

head(res8.6, 4)
```



```{r}
res9 <- as.data.frame(res$laureates$death)   # birthdeath
colnames(res9)[1] = "Birth.Death"  # change column name by colname function
res9 <- subset(res9, select = -c(place)) # drop columns 'place'

head(res9, 4)
```

The Laureates data set contained only 25 values as shown on the embedded table using the `reactable function`.
```{r}
res_cbind = cbind.data.frame(res1, res2, res3, res4, res5, res6, res7, res8, res8.1, res8.2, res8.3, res8.4, res8.5, res8.6,res9)
#head(res_cbind, 5)
```

```{r}
library(reactable)

reactable(res_cbind)
```

## Plot

```{r}
library(ggplot2)
```

## Gender Distribution

### How many female and male laureates recieved nobel prizes based on this dataset? Female(1) and Male(24)
```{r}
country_gender <- res_cbind %>%
  count(res_cbind$Gender)

country_gender
```


```{r}
ggplot(data = res_cbind, aes(x = Birth.PlaceContinent)) +
  stat_count(width = 0.5, aes(fill = Gender)) +
  labs(x = "Continent", y = "Count", 
     title = "Nobel Prizes by Continent and Gender") 
```


## Continent

### On which countinent where these Laureates born?
```{r}
country_plot <- res_cbind %>%
  count(Birth.PlaceContinent)

country_plot
```


```{r}
ggplot(data = res_cbind, aes(x = Birth.PlaceContinent)) +
  stat_count(width = 0.7, aes(fill = Birth.PlaceCountry)) +
  labs(x = "Continent", y = "Count", 
       title = "Nobel Prizes by Continent and Birth Place") +
coord_flip()
```


## Birth Year

### In what year did this dataset of Laureates were born?
```{r}
library("stringr")
birth_year <- str_sub(res_cbind$BirthDate, 1, 4) 

birth_year
```

```{r}
ggplot(data = res_cbind, aes(x = Birth.PlaceContinent, birth_year)) +
  labs(title = "Nobel Prizes by Birth Place Continent and Birth Year") +

geom_point(colour = "blue", size = 3)
```

