---
title: 'CUNY SPS DATA607 Project2: Dataset3'
output:
  html_document:
    df_print: paged
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---
### **Name: Chinedu Onyeka, Date: October 3rd, 2021**
 
##### **Objective:** To determine the percentage of west nile virus disease cases reported to cdc from 1999 to 2019 that turned neuro-invasive by states in the US.  

##### **Data Source:** [CDC](https://www.cdc.gov/westnile/statsmaps/cumMapsData.html) 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Questions: ** The questions for this data set was provided by Victoria McEleney  
West Nile virus disease cases reported to CDC by state of residence, 1999-2019  
West Nile virus neuroinvasive disease cases reported to CDC by state of residence, 1999-2019  
The totals per state & per year are already calculated, but the means are yet to be calculated.  
The year columns could be pivoted longer and the 2 tables could be combined. Percent of positive cases that developed into neuroinvasive disease could be calculated (per year / per state).  

*Load the libraries*
```{r load-library, message=FALSE}
library(tidyverse)
```

*Load the data*
```{r load-file, message=FALSE}
url_westnile_disease <- "https://raw.githubusercontent.com/chinedu2301/DATA607-Data-Acquisition-and-Management/main/West-Nile-virus-disease-cases-by-state_1999-2019-P.csv"
url_westnile_neuro <- "https://raw.githubusercontent.com/chinedu2301/DATA607-Data-Acquisition-and-Management/main/West-Nile-virus-neuroinvasive-disease-cases-by-state_1999-2019-P.csv"

westnile_disease <- read_csv(url_westnile_disease)
westnile_neuro <- read_csv(url_westnile_neuro, skip = 1)
```

*Check the each of the datasets*
```{r}
#View the west nile disease dataset
westnile_disease
```
```{r}
#View the west nile disease cases that turned neuro invasive
westnile_neuro
```
These two data sets are wide and contain a lot of unnecessary rows. Hence, we subset the data sets to contain only the rows for each state and then transform the data sets from wide to long.  

*Subset the data sets*  

```{r}
#slice the west nile disease dataset
westnile_disease <- westnile_disease %>% slice(1:52)
westnile_disease
```
```{r}
#slice the west nile neuro invasive dataset
westnile_neuro <- westnile_neuro %>% slice(1:52)
westnile_neuro
```
*Transform the tables from wide to long*

```{r}
westnile_d <- westnile_disease %>% select(-Total) %>% 
  gather(key = "Year", value = "cases", -State) %>% arrange(State)
westnile_n <- westnile_neuro %>% select(-Total) %>% 
  gather(key = "Year", value = "neuroinvasive_cases", -State) %>% arrange(State)
```

*View the new long data sets*  

West nile disease cases
```{r}
westnile_d
```
West nile neuro invasive cases
```{r}
westnile_n
```
*Combine the two data sets*
```{r}
westnile <- cbind(westnile_d, neuroinvasive_cases = westnile_n$neuroinvasive_cases)
westnile
```
*Find the average (mean) cases for each state: Group by State*

```{r}
westnile_mean <- westnile %>% group_by(State) %>% 
  summarise(Avg_cases = round(mean(cases),0), Avg_neuroinvasive_cases = round(mean(neuroinvasive_cases),0))
westnile_mean
```

*Find percentage of cases that turn neuroinvasive*  
```{r}
westnile_percent_neuro <- westnile_mean %>% 
  mutate(percent_neuroinvasive = round((Avg_neuroinvasive_cases/Avg_cases)*100, 2))
westnile_percent_neuro
```
*Replace the NaN values with NA*
```{r}
westnile_percent_neuro$percent_neuroinvasive[is.nan(westnile_percent_neuro$percent_neuroinvasive)] <- NA
westnile_percent_neuro
```
Construct a bar graph
```{r}
westnile_percent_neuro %>% top_n(30) %>% ggplot(aes(reorder(State, percent_neuroinvasive), percent_neuroinvasive)) + 
  geom_col(fill = "brown") + coord_flip() + xlab("") + labs(title = "Percentage of West Nile virus cases that turned neuroinvasive")
```

*Check summary for the percent_neuroinvasive cases by state*  
```{r}
summary(westnile_percent_neuro$percent_neuroinvasive)
```
*Find the average (mean) cases for each state: Grouping by Year*
```{r}
westnile_month <- westnile %>% group_by(Year) %>% 
  summarise(Avg_cases = round(mean(cases),0), Avg_neuroinvasive_cases = round(mean(neuroinvasive_cases),0))
westnile_percent_month <- westnile_month %>% 
  mutate(percent_neuroinvasive = round((Avg_neuroinvasive_cases/Avg_cases)*100, 2))
#Replace NaN values with NA
westnile_percent_month$percent_neuroinvasive[is.nan(westnile_percent_month$percent_neuroinvasive)] <- NA
westnile_percent_month
```
*Plot a graph of percent neuroinvasive by year*
```{r}
ggplot(westnile_percent_month, aes(Year, percent_neuroinvasive)) + geom_col(fill = "brown") + 
  theme_bw() + labs(title = "Percentage of Neuroinvasive westnile virus cases by year")
```

*Summary statistics for the percentage of west nile virus cases that turned neuroinvasive per year*
```{r}
summary(westnile_percent_month$percent_neuroinvasive)
```


**Conclusion:** *We can see that using the average westnile virus cases from 1999 to 2019 for each state, the median percentage of the westnile virus cases that turn neuroinvasive is about 60% for each state. Also, for each year since 1990, about 60% of westnile virus cases turned neuroinvasive.*






















