
#### **CUNY SPS DATA607 HW2**
#### **Name: Chinedu Onyeka; Date: September 2nd, 2021**

#### This assignment demonstrates ability to connect to a database from R  



Load Libraries
```{r}
library(RODBC) #To access MySQL database 
library(tidyverse)
```



Connect to the MySQL database using odbc connect and a dsn
```{r}
conn <- odbcConnect("MySQL_DSN")
```


Read the reviewers information from MySQL database into R dataframe
```{r}
reviews <- sqlQuery(conn, "SELECT * FROM reviewers")
head(reviews, n=20)
```
```{r}
str(reviews)
```
Check if the dataframe contains any missing values
```{r}
anyNA(reviews)
```
The missing values are in the rating column
```{r}
anyNA(reviews$rating)
```
Missing Values : I omit missing values
```{r}
# avg_rating <- round(mean(reviews$rating, na.rm = TRUE),0)
reviews_new <- reviews %>% na.omit()
```

Check for missingness in the new dataframe
```{r}
anyNA(reviews_new)
```
Check the head of the new dataframe
```{r}
head(reviews_new, n = 20)
```

Check the structure
```{r}
str(reviews_new)
```

Show all 6 videos
```{r}
sqlQuery(conn, "SELECT * FROM videos")
```


