---
title: "CUNY SPS DATA607 Assignment 7"
author: "Chinedu Onyeka"
date: "October 8th, 2021"
output:
  html_document:
    df_print: paged
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### **Question**  

Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more
than one author. For each book, include the title, authors, and two or three other attributes that you find
interesting.  

Take the information that you’ve selected about these three books, and separately create three files which
store the book’s information in HTML (using an html table), XML, and JSON formats (e.g. “books.html”,
“books.xml”, and “books.json”). To help you better understand the different file structures, I’d prefer that you
create each of these files “by hand” unless you’re already very comfortable with the file formats.
Write R code, using your packages of choice, to load the information from each of the three sources into
separate R data frames. Are the three data frames identical?  

Your deliverable is the three source files and the R code. If you can, package your assignment solution up into
an .Rmd file and publish to rpubs.com. [This will also require finding a way to make your three text files
accessible from the web].  


### *Solution*

```{r load-libraries, message=FALSE}
# load libraries
library("tidyverse")
library("xml2")
library("jsonlite")
library("rvest")
library("XML")
library("DT")
```




#### **HTML**  

```{r}
url_html<-"https://raw.githubusercontent.com/chinedu2301/DATA607-Data-Acquisition-and-Management/main/books.html"

# read the html file and select the required nodes
book_html <- read_html(url_html) %>% html_nodes("table")

# convert to table
book_html <- html_table(book_html, header = TRUE)[[1]] %>% as.data.frame() %>% datatable()
  
# show the table
book_html
```

#### **XML**  

```{r}
url_xml<-"https://raw.githubusercontent.com/chinedu2301/DATA607-Data-Acquisition-and-Management/main/books.xml"

# read the xml file
book_xml <- read_xml(url_xml)

# parse the xml file and convert into a table
book_xml <- xmlParse(book_xml) %>% xmlToDataFrame() %>% datatable()

# show the table
book_xml
```

#### **JSON**  

```{r}
url_json<-"https://raw.githubusercontent.com/chinedu2301/DATA607-Data-Acquisition-and-Management/main/books.json"

# read the json file and convert to table
book_json <- fromJSON(url_json) %>% as.data.frame() %>% datatable()

# show the table
book_json
```
*How they differ: *  
Looking at the three tables, they look similar/identical to me. However, I will let r to determine if all three are identical:  

```{r}

html_vs_xml <- identical(book_html, book_xml)
html_vs_json <- identical(book_html, book_json)
xml_vs_json <- identical(book_xml, book_json)
paste0("Are the html and xml books table identical? ", " Answer: ", html_vs_xml)
paste0("Are the html and json books table identical? ", " Answer: ", html_vs_json)
paste0("Are the xml and json books table identical? ", " Answer: ", xml_vs_json)
```
Even though the output looks similar for all three tables, they are not identical to one another. They have different internal structure and R handles them differently to produce an output that looks similar.






