---
title: "CUNY SPS DATA607 Assignment 9"
author: "Chinedu Onyeka"
date: "October 23rd, 2021"
output:
  html_document:
    df_print: paged
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

**Brief Description**  
*This assignment uses the New York Times web API to obtain information about top science stories from NYT.*  

**Approach**  
*To access the NYT web API, a developer account is needed. So, I created a developer account on the NYT website after which I requested an API key for their top science stories. To use the API to access the data, I used the GET() function from the `httr`package to make a GET request to the NYT server. After a successful GET request with a status code of `200` which signifies that the request was successful, I parsed the content of the JSON response, used the fromJSON() from `jsonlite` package to read the JSON file, flattened the output and then converted the results to dataframe.*  


*Load the required libraries*
```{r libraries, message = FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(jsonlite)
library(rlist)
```

*Make API call*

```{r api-call}
nyt_top_stories <- GET("https://api.nytimes.com/svc/topstories/v2/science.json?api-key=zYpgBL2MhyOOlSeLaDAKfQJQo2BeQZlz")
```

*Inspect the status of the API requests*
```{r inspect-status}
# Check the status of the api call
nyt_top_stories
```

*Since the status code is 200, it means it is OK.*  

```{r parse}
# Parse the content of the response from the API call
nyt_parse <- content(nyt_top_stories, "parse")
nyt_parse_results <- nyt_parse$results
```

```{r flatten}
nyt_top <- fromJSON(rawToChar(nyt_top_stories$content))
#check the names of the ny_top
names(nyt_top)
```

```{r}
# Top stories data frame
nyt_top_flat <- flatten(nyt_top$results) 
nyt_top_flat <- as.data.frame(nyt_top_flat)
nyt_top_flat
```
*Some columns came in as a list and some as a dataframe.*
```{r multimedia}
#Take a look at the multimedia dataframe which is a column in the 
as.data.frame(nyt_top_flat$multimedia)
```



```{r}
# Check the elements of des_facet column that came in as a list
nyt_top_flat$des_facet
```

```{r}
# Check the elements of the org_facet column that also came in as a list
nyt_top_flat$org_facet
```
**Conclusion**  
*The response from NYT server was a nested JSON file. When converted to dataframe, I found out that some columns contained lists while another column contained a dataframe. So a dataframe is a column in another dataframe. To access the content of that dataframe, I obtained the column containing a dataframe and displayed its results.*  
*In conclusion, when a highly nested JSON file is read into a dataframe, some columns of the dataframe may be a list while some may be a dataframe. JSON files are very much effective in storing large amounts of unstructured data. No wonder most if not all responses of API calls in most websites returns a JSON file.* 









